---
title: "Stats 20, F23 -- Homework 4"
author: "Daren Sathasivam -- 306229580"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

## Question 1

### (a)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 1}
my_t <- function(x) {
  # Initializers
  rows <- nrow(x)
  cols <- ncol(x)
  t_mat <- matrix(0, cols, rows)
  # Check for vector and convert if not
  if(is.vector(x)) {
    x <- matrix(x, ncol = 1)
  }
  # For-loop through 2d matrix
  for (i in 1:rows) {
    for (j in 1:cols) {
      t_mat[j, i] <- x[i, j] # Swap locations i.e. [1, 2] --> [2, 1]
    }
  }
  # Transposed back to vector
  if(is.vector(x)) {
    t_mat <- as.vector(t_mat)
  }

  t_mat
}

x <- matrix(1:9, nrow = 3, ncol = 3)
x
my_t(x)
t(x)
```


## Question 2

### (a)
```{r Question 2a}
my_cor <- function(x, y, use = "everything") {
  # x and y numeric check
  if(!is.numeric(x) || !is.numeric(y)) {
    stop("Both x and y must be numeric!")
  }

  # Remove the missing value using "use"
  if(use == "everything") {
    n <- length(x)
    numerator <- sum((x - mean(x)) * (y - mean(y)))
    sd_x <- sum((x - mean(x))^2)
    sd_y <- sum((y - mean(y))^2)
    cor_cof <- numerator / sqrt(sd_x * sd_y)
  } else if (use == "pairwise.complete.obs") {
      # Get valid indices
      na_idx <- !is.na(x) & !is.na(y)
      x <- x[na_idx]
      y <- y[na_idx]
      my_cor(x, y, use = "everything")
  } else {
    stop("Invalid 'use' argument!")
  }
  cor_cof
}
x <- c(1, 2, 3, 4, 5)
y <- c(6, 7, 8, 9, 2)
cor(x, y)
my_cor(x, y)

z <- c(NA, 2, 4)
a <- c(NA, 3, 5)
cor(z, a)
my_cor(z, a)
```
<!-- ENTER YOUR EXPLANATION BELOW, FEEL FREE TO EDIT THE ABOVE CODE CHUNK -->
### (b)
```{r}
linreg <- function(x, y, use = "everything") {
  b <- (my_cor(x, y) * sd(y)) / sd(x)
  a <- mean(y) - (b * mean(x))
  c <- c(a, b)
  c
}
x <- c(1, 2, 3, 4, 5)
y <- c(6, 7, 8, 9, 2)
linreg(x, y)
```
### (c)
```{r}
height <- c(61, 62, 63, 64, 66, 68)
weight <- c(104, 110, 125, 141, 160, 170)
# Find equation of regression line using linreg
ab <- linreg(height, weight)
a <- ab[1]
b <- ab[2]

message("Weight = ", round(a, 2), " + ", round(b, 2), " * ", "height")
weight <- round(a, 2) + (round(b, 2) * height)
weight
```
### (d)
```{r}
linreg_mat <- function(x, y) {
  X <- cbind(rep(1, length(x)), x)
  if (length(x) == length(y)) {
    solve(t(X) %*% X) %*% t(X) %*% y
  } else {
    warning("Length of x and y are not the same!")
  }
}
```
### (e)
```{r}
ab_mat <- linreg_mat(height, weight)
a_mat <- ab_mat[1]
b_mat <- ab_mat[2]
a_mat
b_mat

message("Using 'linreg_mat': Weight = ", round(a_mat, 2), " + ", round(b_mat, 2), " * ", "height")
```
### (f)
Weight = -510.65 + 10.09 * height\
The slope coefficient of 10.09 represents that for every inch difference, there should be roughly a 10.09 pound difference.


## Question 3

### (a)
```{r Question 3a}
'%m%' <- function(A, B) {
  if (ncol(A) != nrow(B)) {
    stop("Dimensions are not compatible!")
  }
  # Initializer
  result_mat <- matrix(0, nrow = nrow(A), ncol = ncol(B))
  # Matrix Multiplication
  for (i in 1:nrow(A)) {
    for (j in 1:ncol(B)) {
      # Sum of A_ith and B_jth 
      result_mat[i, j] <- sum(A[i, ] * B[, j])
    }
  }
  result_mat
}
```

### (b)
<!-- ENTER YOUR ANSWER BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 3b}
X <- matrix(c(6, 3, 5, 2, 4, 1), nrow = 2, ncol = 3)
X
Y <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3, ncol = 3)
Y

x_times_y <- X %m% Y
x_times_y
X %*% Y
```

## Question 4

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 4a}
'%^%' <- function(A, k) {
  if (!is.numeric(k) || k < 0) {
    stop("kth value should be a nonnegative numeric value!")
  }
  
  # When kth is 0, out put a diag for the matrix
  if (k == 0) {
    result_mat <- diag(nrow(A))
  } else {
    result_mat <- A
    # Multiply the A matrix k times itself
    for (i in 2:k) {
      result_mat <- result_mat %m% A
    }
  }
  result_mat
}
```

### (b)
<!-- USE THE CODE CHUNK BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 4b}
Z = matrix(c(0.2, 0.6, 0.4, 0.7, 0.2, 0.1, 0.1, 0.2, 0.5), nrow = 3, ncol = 3)
Z

Z_0 <- Z %^% 0
Z_0
Z_5 <- Z %^% 5
Z_5
Z_50 <- Z %^% 50
Z_50
Z_500 <- Z %^% 500
Z_500
```


## Question 5

<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5a}
my_which <- function(x, arr.ind = FALSE) {
  if (!is.logical(x) && !is.numeric(x)) {
    stop("Input must be a logical/numeric vector or matrix!")
  }

  if (is.matrix(x)) {
    # Index initializers
    get_idx <- array(1:(nrow(x) * ncol(x)), dim = c(nrow(x), ncol(x)))
    row_idx <- rep(1:nrow(x), each = ncol(x))
    col_idx <- rep(1:ncol(x), times = nrow(x))
    # If arr.ind = TRUE output with row and col indices
    if (arr.ind) {
      # Unsure of why row_idx and col_idx are swapped when arr.ind
      get_idx <- cbind(row = col_idx[x], col = row_idx[x])
    } else {
      get_idx <- seq_along(x)
      get_idx <- get_idx[x]
    }
  } else if (is.vector(x)) { # For vector
    get_idx <- seq_along(x)
    if (arr.ind) {
      get_idx <- cbind(row = get_idx[x])
    } else {
      get_idx <- get_idx[x]
    }
  }
  get_idx
}

x <- matrix(c(TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE, TRUE, FALSE), nrow = 3, ncol = 3)
x

my_which(x)
my_which(x, arr.ind = TRUE)
which(x, arr.ind = TRUE)

# -- Not accurate for numeric matrtix -- #
# y <- matrix(1:6, nrow = 2, ncol = 3)
# y
# my_which(y %% 2 == 0, arr.ind = TRUE)
# my_which(y %% 2 == 0)
# which(y %% 2 == 0, arr.ind = TRUE)
```

## Question 6
### (a)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 6}
load("mlb.RData")
# i.
tapply(hit, team, max)

# ii.
table(team)

# iii.
table(team[hr > 0])

```
### (b)
```{r}
# i.
tapply(hit[ab >= 100] / ab[ab >= 100], team[ab >= 100], max)

# ii.
tapply(hit / ab, team, mean, na.rm = TRUE)

```

### (c)
```{r}
tapply(hr, list(league, pos), mean) # [max(diff(team))]
```
The position with the largest difference between leagues is rightfield between the American League and National League.

### (d)
```{r}
tapply(1:length(team), pos, median)
```
 


## Question 7

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 7a}
my_row <- function(x) {
  if (!is.matrix(x)) {
    stop("Input must be a matrix!")
  }
  matrix(1:nrow(x), nrow = nrow(x), ncol = ncol(x))
}
x <- matrix(1:9, nrow = 3, ncol = 3)
x

my_row(x)
row(x)
```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 7b}
my_col <- function(x) {
  if (!is.matrix(x)) {
    stop("Input must be a matrix!")
  }
  matrix(1:ncol(x), nrow = nrow(x), ncol = ncol(x), byrow = TRUE)
}
x <- matrix(1:9, nrow = 3, ncol = 3)
my_col(x)
col(x)
```

### (c)
<!-- CREATE THE FUNCTIONs IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTIONs AND THEIR INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 7c}
my_diag <- function(x, nrow = NULL, ncol = NULL) {
  # Initialize empty matrix
  diag_mat <- matrix(0, nrow = x, ncol = x)
  
  if (missing(nrow) && missing(ncol)) { # For when nrow and ncol default to NULL
    # Scalars case
    if (length(x) == 1) {
      if (x == 0) {
        diag_mat <- matrix(0, nrow = 0, ncol = 0)
      } else {
        for (i in 1:x) {
          diag_mat[i, i] <- 1
        }
      }
      
    } else if (is.null(dim(x))) { #Vectors case
      diag_mat <- matrix(0, nrow = length(x), ncol = length(x)) 
        for (i in 1:length(x)) {
          diag_mat[i, i] <- x[i]
        }
    } else { # Matrices case
      # Initialize empty vector
      diag_mat <- numeric(0) 
      for (i in 1:dim(x)[1]) {
        # Combine empty vec with values at the [i,i] indices of the given matrix
        diag_mat <- c(diag_mat, x[i, i])
      }
    }
  } else { # For when nrow and ncol are specified
    diag_mat <- matrix(0, nrow, ncol)
      # Iterate through the length of the shorter dim
      for (i in seq_len(min(nrow, ncol))) {
        diag_mat[i, i] <- x[(i - 1) %% length(x) + 1]
      }
  }
  diag_mat
}

my_diag(0)
diag(0)
my_diag(3)
diag(3)
my_diag(1:3)
diag(1:3)
my_diag(matrix(1:9, 3, 3))
diag(matrix(1:9, 3, 3))
my_diag(matrix(1:3), 4, 5)
# diag(matrix(1:3), 4, 5)
```

## Question 8
```{r}
load("votes.RData")
```

### (a)
<!-- CREATE THE FUNCTIONs IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTIONs AND THEIR INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
Pseudocode:\
Function : tally_rcv \
\
Input: \
- A matrix of ranked choice votes (votes_mat)\
\
Ouput:\
- A matrix of results (results_mat) \ 
\
Algorithm: \
- Initialize the results matrix with the appropriate dimensions. \
- Extract the number of choices/candidates (num_candidates) from votes_mat. \
- Extract the number of total votes (total_voters) from votes_mat. \
- Initialize eliminated_candidates list as an empty vector 'numeric(0)'. \
- Initialize "round" number as 1. \
- While loop to repeat until a winner is determined. \
  - Initialize an array to store the count of votes for each candidate  (vote_count).
  - Count the votes for each remaining candidate: \
    - Iterate through each voter: \
      - Identify the candidate with the highest-ranked choice who has not been eliminated. \
      - Increment the count for the identified candidate. \
  - Identify the candidate(s) with the fewest votes (lowest_votes). \
  - If there is a tie for last place: \
    - In the first round, eliminate the tied candidate who comes last alphabetically using their first name. \
    - In subsequent rounds: \
      - If the sum of the tied candidates' vote is less than the next lowest candidates' cotes: \
        - Eliminate both of the tied candidates. \
      - Else, eliminate the tied candidate with the least votes in the previous rounds. \
  - Update the results_mat with the current round's winners and eliminated candidates. \
  - If any candidate has more than half of the total votes, declare them the winner and exit the while loop. \
  - Increment the round number (round). \
- Return results_mat \

### (b)
<!-- USE THE CODE CHUNK BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 8b}
tally_rcv <- function(votes_mat) {
  
}
```

### (c)
```{r}
# knitr::kable(tally_rcv(votes_mat, format = "html"))
```

